Explore expression of marker genes for tumor cell states

### 2025-01-13 analyses/cell-type-ewings/exploratory_analysis/06-tumor-cell-state-assignment.Rmd
respond to reviews

In this notebook we look at expression of potential markers for tumor cell states in two libraries from `SCPCP000015`: `SCPCL000822` and `SCPCL000824`. 
The expression of specific marker genes are then used to classify tumor cells into three different tumor cell states: `EWS-low`, `EWS-mid`, and `EWS-high`.

@@

`SCPCL000822`: 
`EWS-FLI1` low  - 2
`EWS-FLI1` middle - 1 & 4
`EWS-FLI1` high - 3

`SCPCL000824` : 
`EWS-FLI1` low - 1
`EWS-FLI1` middle - 6 
`EWS-FLI1` high - 3, 4, 5, 7
`EWS-FLI1` high proliferative - 2

@@

- There is a clear division between cells that are `EWS-FLI` high and `EWS-FLI1` low. 
- Most tumor cells in both samples are `EWS-FLI` high, which is consistent with the literature. 
- There may be a class of cells that live in between high and low, but if we want to define those we probably need to be more stringent with our approach and identify a mechanism to bin cells based on expression of both `EWS-FLI1` activated and repressed genes, which is outside the scope of this notebook.

>>>>

In this notebook we look at expression of potential markers for tumor cell states in two libraries from `SCPCP000015`: `SCPCL000822` and `SCPCL000824`. 
The expression of specific marker genes are then used to classify tumor cells into three different tumor cell states: `EWS-low`, `EWS-mid`, `EWS-high`, and `EWS-high-proliferative`.

@@

`SCPCL000822`:

`EWS-FLI1` low  - 2
`EWS-FLI1` middle - 1 & 4
`EWS-FLI1` high - 3

`SCPCL000824` :

`EWS-FLI1` low - 1
`EWS-FLI1` middle - 6 
`EWS-FLI1` high - 3, 4, 5, 7
`EWS-FLI1` high proliferative - 2

@@

- There is a clear division between cells that are `EWS-FLI` high and `EWS-FLI1` low. 
  - We also see a set of cells that are both "proliferative" and `EWS-FLI` high. 
- Most tumor cells in both samples are `EWS-FLI` high, which is consistent with the literature. 
- There may be a class of cells that live in between high and low, but if we want to define those we probably need to be more stringent with our approach and identify a mechanism to bin cells based on expression of both `EWS-FLI1` activated and repressed genes, which is outside the scope of this notebook.
